Calculating impedances for CMA and CA areas from the GSS surveys

### 2024-11-12 data-raw/01-preprocessing-source-files.Rmd
Manuscript

```{r}
walking_2015 <- walking_2015 |>
  filter(origin == 300 | 
           origin == 301 | 
           origin == 302 | 
           origin == 303 | 
           origin == 304 |
           origin == 305 | 
           origin == 306 | 
           origin == 307 |
           origin == 308 |
           origin == 309 | 
           origin == 310 |
           origin == 311) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 302 ~ "Business",
                     origin == 303 ~ "Other's home",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic"),
         orig_label = factor(orig_label))
```

@@

```{r}
walking_2010 <- walking_2010 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 7 ~ "Grocery store, other stores or mall",                     
                     origin == 8 ~ "Work or school"),
         orig_label = factor(orig_label))
```

>>>>

```{r}
walking_2015 <- walking_2015 |>
  filter(origin == 300 | 
           origin == 301 | 
           origin == 302 | 
           origin == 303 | 
           origin == 304 |
           origin == 305 | 
           origin == 306 | 
           origin == 307 |
           origin == 308 |
           origin == 309 | 
           origin == 310 |
           origin == 311) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 302 ~ "Business",
                     origin == 303 ~ "Other's home",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic"),
         orig_label = factor(orig_label))
```

@@

```{r}
walking_2010 <- walking_2010 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 7 ~ "Grocery store, other stores or mall",                     
                     origin == 8 ~ "Work or school",
                     origin == 9 ~ "Outdoors",
                     origin == 10 ~ "Library, museum or theatre"),
         orig_label = factor(orig_label))
```

### 2024-11-12 data-raw/02-calculating-impedances.Rmd
Manuscript

```{r}
gss_episodes <- gss_episodes %>%
  group_by(MODE, YEAR, destination) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  filter(count > 1)

@@

Similar as we did before, the next chunk selects the best impedance function, using the `test_distributions_weighted` and `lowest_aic` functions for each destination, transportation mode and year. The name pattern is as follows: `f_year_mode_dest` + `destination`, which means that `f_1986_Walking_dest1` displays the best function for the year 1986, walking mode, destination 1 (Home).
/000
```{r}

  for(destino in unique(gss_episodes$destination)){

    for(mode in unique(gss_episodes[gss_episodes$YEAR == year, ]$MODE)){

        gss_subset <-  gss_episodes[gss_episodes$destination == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode, ]

        if(nrow(gss_subset) > 0){
          print(paste("Calculating the best function for ", 
                    mode, ", destination ",  
                    gss_subset$dest_label[1],
                    ", for the ", year, " GSS cicle"))

@@

  for(destino in unique(gss_episodes$destination)){

    for(mode in unique(gss_episodes[gss_episodes$YEAR == year, ]$MODE)){

        gss_subset <-  gss_episodes[gss_episodes$destination == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode ,
                            ]

@@

                    print(paste("Obtaining the values for ", 
                    mode, ", destination ",  
                    gss_subset$dest_label[1],
                    ", for the ", year, " GSS cicle"))

@@

        gss_episodes[gss_episodes$destination == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode ,
                            c("f_name","f","est_1","est_2")] <- df_f[,c("f_name","f","est_1","est_2")]

        gss_episodes[gss_episodes$destination == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode ,
                            c("loglik", "aic", "bic")] <- list(get(current_function)$loglik, get(current_function)$aic, get(current_function)$bic)
        }
      }
  }
}

>>>>

```{r}
gss_episodes <- gss_episodes %>%
  group_by(MODE, YEAR, dest_label) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  filter(count > 1)

@@

Similar as we did before, the next chunk selects the best impedance function, using the `test_distributions_weighted` and `lowest_aic` functions for each destination, transportation mode and year. The name pattern is as follows: `f_year_mode_dest` + `dest_label`, which means that `f_1986_Walking_destHome` displays the best function for the year 1986, walking mode, Home destination.
/000
```{r}
for(year in unique(gss_episodes$YEAR)){

  for(destino in unique(gss_episodes[gss_episodes$YEAR == year, ]$dest_label)){

    for(mode in unique(gss_episodes[gss_episodes$YEAR == year & gss_episodes$dest_label == destino, ]$MODE)){

        gss_subset <-  gss_episodes[gss_episodes$dest_label == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode, ]

        if(nrow(gss_subset) > 0){
          print(paste("Calculating the best function for ", 
                    mode, ", destination ",  
                    destino,
                    ", for the ", year, " GSS cicle"))

@@

  for(destino in unique(gss_episodes[gss_episodes$YEAR == year, ]$dest_label)){

    for(mode in unique(gss_episodes[gss_episodes$YEAR == year & gss_episodes$dest_label == destino, ]$MODE)){

        gss_subset <-  gss_episodes[gss_episodes$dest_label == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode ,
                            ]

@@

                    print(paste("Obtaining the values for ", 
                    mode, ", destination ",  
                    destino,
                    ", for the ", year, " GSS cicle"))

@@

        gss_episodes[gss_episodes$dest_label == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode ,
                            c("f_name","f","est_1","est_2")] <- df_f[,c("f_name","f","est_1","est_2")]

        gss_episodes[gss_episodes$dest_label == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode ,
                            c("loglik", "aic", "bic")] <- list(get(current_function)$loglik, get(current_function)$aic, get(current_function)$bic)
        }
      }
  }
}
